Space-time plots

### 2016-04-05 Rmd/space-time.Rmd
1. Removed commented out code for old resampling with perturbation approach. 2. Cleaned up and compiled html output of space-time

# Space versus time analysis
```{r, eval = FALSE}
library(ecoscales)
library(readxl)
library(stringdist)
library(maptools)

@@

dat <- dat[, names(dat)[1:20], with = FALSE]
unique(dat$DOI)  # 138
```

## Data prep

### Merge full and calibration datasets
```{r, eval = FALSE}
dat2 <- copy(dat[, .(plot_res, n_sites, sampled_area, samp_duration,
                     t_btwn_samp, study_duration)])

@@

```{r, eval=FALSE}
# apply fixes (mostly to account for observer omissions)
# 1. if samp_duration != study_duration and t_btwn_samp == 0, set 
#    study_duration == samp_duration
names(datf)
datf[t_btwn_samp == 0 & (study_duration != samp_duration), ]  
dat[t_btwn_samp == 0 & (study_duration != samp_duration), c(1, 3, 8:13), with=F]  
# function to set t_btwn_samp to 365 * 100000 to indicate once-off studies 
# when t_btwn_samp == 0.  
to_infin_byond <- function(tbtwn) ifelse(tbtwn == 0, 365 * 10000, tbtwn)

@@

### Bootstraps

### With uncertainties

@@

```{r, eval=FALSE}
p <- c(0.79, 0.99, 0.53, 1.22)

@@

# old function, retained here for easy recollection of first approach
# full function for resampling
# resamp_func <- function(dat, p, iter) {
#   bout <- lapply(1:iter, function(x) {  # x <- 1
#     if((x / 100) %in% 1:100) print(x)
#     # amount to perturb
#     a <- runif(1, min = 1 - p, max = 1 + p)  # plot_res perturb
#     b <- runif(1, min = 1 - p, max = 1 + p)  # n_sites perturb
#     d <- runif(1, min = 1 - p, max = 1 + p)  # samp_duration perturb
#     e <- runif(1, min = 1 - p, max = 1 + p)  # t_btwn_samp perturb
#     
#     if(length(p) < 4) stop("p must be length of 4", call. = FALSE)
# 
#     
#     # spatial variables
#     dnew <- copy(dat[, .(plot_res, n_sites, sampled_area, samp_duration,
#                          t_btwn_samp, study_duration)])
#     dnew[which(senst$res == 1), plot_res := plot_res * a]
#     dnew[which(senst$n == 1), n_sites := n_sites * b]
#     dnew[, sampled_area := (plot_res * n_sites) / 10000]
# 
#     # temporal variables
#     dnew[which(senst$sampd == 1), samp_duration := samp_duration * d]
#     dnew[which(senst$tbtwn == 1), t_btwn_samp := t_btwn_samp * e]
#     dnew[, study_duration := {
#       recalc_duration(study_duration, samp_duration, t_btwn_samp)
#     }]
#     dnew[, t_btwn_samp := to_infin_byond(t_btwn_samp)]  # set high
#   })
#   return(bout)
# }

# perturbed samples
# p10 <- resamp_func(dat, 0.1, 1000)  # 10% perturb

@@

### Without uncertainties

@@

### Perturbed boot-strap
```{r, eval = FALSE}
bootp_dt <- rbindlist(bootperturb)
resdat <- data.table(x = log10(bootp_dt$t_btwn_samp), 
                      y = log10(bootp_dt$plot_res))
resdat$n <- rep(1, nrow(resdat))
rx <- range(resdat$x)
rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
# rx[1] <- -5
rx[2] <- 6.56
resdat[x < rx[1], x := rx[1]]
resdat[x > rx[2], x := rx[2]]
# rx[2] <- 7.6
ry <- range(aaxis1$logres)
ry[1] <- -5  # set limits to 0.01 cm^2
resdat[y < ry[1], y := ry[1]]
resdat[y > ry[2], y := ry[2]]
coordinates(resdat) <- ~x + y

@@

### Ordinary bootstrap
```{r, eval = FALSE}
bootdt <- rbindlist(bootsamp)
resdat2 <- data.table(x = log10(bootdt$t_btwn_samp), y = log10(bootdt$plot_res))
resdat2$n <- rep(1, nrow(resdat2))
rx <- range(resdat2$x)
rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
rx[2] <- 6.56
resdat2[x < rx[1], x := rx[1]]
resdat2[x > rx[2], x := rx[2]]
ry <- range(aaxis1$logres)
ry[1] <- -5  # set limits to 0.01 cm^2
resdat2[y < ry[1], y := ry[1]]
resdat2[y > ry[2], y := ry[2]]
coordinates(resdat2) <- ~x + y

@@

### On primary dataset only
```{r, eval = FALSE}
resdat3 <- data.table(x = log10(to_infin_byond(datf$t_btwn_samp)), 
                      y = log10(datf$plot_res))
resdat3$n <- rep(1, nrow(resdat3))
rx <- range(resdat3$x)
rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
rx[2] <- 6.56
resdat3[x < rx[1], x := rx[1]]
resdat3[x > rx[2], x := rx[2]]
ry <- range(aaxis1$logres)
ry[1] <- -5  # set limits to 0.01 cm^2
resdat3[y < ry[1], y := ry[1]]
resdat3[y > ry[2], y := ry[2]]
coordinates(resdat3) <- ~x + y

@@

```{r, eval = FALSE}
svars <- c("plot_res", "n_sites", "sampled_area", "samp_duration",
           "t_btwn_samp", "study_duration")

sens <- gsub("\\\\", ";", dat$sensitivity)
sens <- gsub(", ", ";", gsub("; ", ";", sens))
sens <- tolower(sens)
sensl <- strsplit(sens, ";")
unique(unlist(sensl))

@@

for(i in 1:nrow(corrtab)) {
  sens2 <- gsub(corrtab[i, 1], corrtab[i, 2], sens2)
}
# gsub("[a-z]( )[a-z]", "_", sens[[228]])
# sens[[228]]
# sens2[is.na(sens2)]
sens2[is.na(sens2)] <- "none"
sensl2 <- lapply(1:length(sens2), function(x) strsplit(sens2[x], ";")[[1]])
# unique(unlist(sensl))
unique(unlist(sensl2))

@@

# percentage of records that are uncertain
colSums(senst) / nrow(senst) * 100

@@

### Varying kernel sizes on resolution versus extent
```{r,eval = FALSE}
# resolution data
# resdat <- data.table(x = log10(datf$t_btwn_samp), y = log10(datf$plot_res))
# resdat$n <- rep(1, nrow(resdat))
# coordinates(resdat) <- ~x + y
# rx <- range(resdat$x)
# # rx <- range(aaxis$logarea)
# rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
# # rx[1] <- -6.5  # set temporal interval to -6.5 to have null returns centered
# # ry <- range(resdat$y)
# ry <- range(aaxis$logarea)
# ry <- round(ry / 0.5) * 0.5  # round to nearest 0.5
# 
# # extent data
# # fix area rounded to zero in Jon's plots
# extdat <- data.table("x" = log10(dat$study_duration), 
#                      "y" = log10(dat$sampled_area)) 
# # extdat[is.infinite(x), ]; extdat[is.infinite(y), ]
# 
# extdat$n <- rep(1, nrow(extdat))
# coordinates(extdat) <- ~x + y
# # ex <- range(extdat$x)
# ex <- range(taxis2$logdays)
# ex <- round(ex / 0.5) * 0.5  # round to nearest 0.5
# # ex[1] <- -5
# ey <- range(extdat$y)
# ey <- round(ey / 0.5) * 0.5  # round to nearest 0.5

# density rasters
# vary the kernel size about them
kernsl <- lapply(c(0.4, 0.6, 0.8, 1), function(x) {
  rr <- kdensity(rx[1], rx[2], ry[1], ry[2], 0.1, resdat3, x)
  er <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat3, x)
  kerns <- list("res" = rr, "ext" = er)
})

# As separate plots
pdf("paper/figures/res_v_extent_ksize.pdf", width = 7, height = 12, )
par(mfrow = c(4, 2), oma = c(2, 2, 0, 0), mar = c(4, 4, 4, 4))
lapply(kernsl, function(x) {
  # x <- kernsl[[1]]
  image(x$res, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
        ylab = "")#, ylim = range(aaxis1$logarea))
  axis(1, at = taxis1$logdays[-c(5, 9, 10)], labels = tlab1[-c(5, 9, 10)], 
       las = 2, tcl = -0.2, cex.axis = cxa)
  axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
       cex.axis = cxa)  
  aargs <- list(mgp = c(3, 0.25, 0), at = brklabs, labels = brklabs, 
                cex.axis = cxa, tcl = -0.1)
  plot(x$res, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))

  image(x$ext, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
        ylab = "")
  axis(1, at = taxis2$logdays, labels = taxis2$label, las = 2)
  axis(2, at = aaxis2$logarea, labels = alab2, las = 2)
  aargs <- list(mgp = c(3, 0.25, 0), at = brklabs2, labels = brklabs2, 
                cex.axis = cxa, tcl = -0.1)
  plot(x$ext, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
       legend.shrink = 0.9, breaks = brks2, col = rev(terrain.colors(30)))
})
par(xpd = NA)
xx <- grconvertX(0.5, from = "ndc", to = "inches")
ii <- c(0.4, 0.6, 0.8, 1)
yy <- c(0.95, 0.75, 0.5, 0.25)
# yy <- c(-1, -5, -10, -20)
for(i in 1:length(ii)) {
  y <- grconvertY(yy[i], from = "ndc", to = "inches")
  # text(x = xx, y = y, labels = paste("kernel =", ii[i]))
  # par(xpd = NA)
  # mtext(text = paste("kernel =", ii[i]), side = 3, line = yy[i])
}
dev.off()

>>>>

# Space versus time analysis
```{r, message = FALSE, warning=FALSE, results='hold'}
library(ecoscales)
library(readxl)
library(stringdist)
library(maptools)

@@

dat <- dat[, names(dat)[1:20], with = FALSE]
```

`r length(unique(dat$DOI))` papers  main analysis

## Data prep
### Merge full and calibration datasets
```{r, message=FALSE, warning=FALSE, results='hide'}
dat2 <- copy(dat[, .(plot_res, n_sites, sampled_area, samp_duration,
                     t_btwn_samp, study_duration)])

@@

```{r, message=FALSE, warning=FALSE, results='hide'}
# apply fixes (mostly to account for observer omissions)
# 1. if samp_duration != study_duration and t_btwn_samp == 0, set 
#    study_duration == samp_duration
names(datf)
datf[t_btwn_samp == 0 & (study_duration != samp_duration), ]  
dat[t_btwn_samp == 0 & (study_duration != samp_duration), c(1, 3, 8:13), with=F]  
# function to set t_btwn_samp to 365 * 100000 to indicate once-off studies 
# when t_btwn_samp == 0.  
to_infin_byond <- function(tbtwn) ifelse(tbtwn == 0, 365 * 10000, tbtwn)

@@

## Analyses

### Resampling with uncertainties

@@

```{r, eval=FALSE}
p <- c(0.79, 0.99, 0.53, 1.22)  # from calibration.Rmd

@@

@@

### Bootstrapping

@@

### Perturbed set
```{r, eval = FALSE}
bootp_dt <- rbindlist(bootperturb)
resdat <- data.table(x = log10(bootp_dt$t_btwn_samp), 
                      y = log10(bootp_dt$plot_res))
resdat$n <- rep(1, nrow(resdat))
rx <- range(resdat$x)
rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
# rx[1] <- -5
rx[2] <- 6.56
resdat[x < rx[1], x := rx[1]]
resdat[x > rx[2], x := rx[2]]
# rx[2] <- 7.6
ry <- range(aaxis1$logres)
ry[1] <- -5  # set limits to 0.01 cm^2
resdat[y < ry[1], y := ry[1]]
resdat[y > ry[2], y := ry[2]]
coordinates(resdat) <- ~x + y

@@

### Bootstrap
```{r, eval = FALSE}
bootdt <- rbindlist(bootsamp)
resdat2 <- data.table(x = log10(bootdt$t_btwn_samp), y = log10(bootdt$plot_res))
resdat2$n <- rep(1, nrow(resdat2))
rx <- range(resdat2$x)
rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
rx[2] <- 6.56
resdat2[x < rx[1], x := rx[1]]
resdat2[x > rx[2], x := rx[2]]
ry <- range(aaxis1$logres)
ry[1] <- -5  # set limits to 0.01 cm^2
resdat2[y < ry[1], y := ry[1]]
resdat2[y > ry[2], y := ry[2]]
coordinates(resdat2) <- ~x + y

@@

### Primary dataset (main figure)
```{r, eval = FALSE}
resdat3 <- data.table(x = log10(to_infin_byond(datf$t_btwn_samp)), 
                      y = log10(datf$plot_res))
resdat3$n <- rep(1, nrow(resdat3))
rx <- range(resdat3$x)
rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
rx[2] <- 6.56
resdat3[x < rx[1], x := rx[1]]
resdat3[x > rx[2], x := rx[2]]
ry <- range(aaxis1$logres)
ry[1] <- -5  # set limits to 0.01 cm^2
resdat3[y < ry[1], y := ry[1]]
resdat3[y > ry[2], y := ry[2]]
coordinates(resdat3) <- ~x + y

@@

```{r}
svars <- c("plot_res", "n_sites", "sampled_area", "samp_duration",
           "t_btwn_samp", "study_duration")

sens <- gsub("\\\\", ";", dat$sensitivity)
sens <- gsub(", ", ";", gsub("; ", ";", sens))
sens <- tolower(sens)
sensl <- strsplit(sens, ";")
# unique(unlist(sensl))

@@

for(i in 1:nrow(corrtab)) {
  sens2 <- gsub(corrtab[i, 1], corrtab[i, 2], sens2)
}
# gsub("[a-z]( )[a-z]", "_", sens[[228]])
# sens[[228]]
# sens2[is.na(sens2)]
sens2[is.na(sens2)] <- "none"
sensl2 <- lapply(1:length(sens2), function(x) strsplit(sens2[x], ";")[[1]])
# unique(unlist(sensl))
# unique(unlist(sensl2))

@@

# percentage of records that are uncertain
round(colSums(senst) / nrow(senst) * 100, 1)

@@

### Varying kernel sizes
```{r,eval = FALSE}
# resolution data
# resdat <- data.table(x = log10(datf$t_btwn_samp), y = log10(datf$plot_res))
# resdat$n <- rep(1, nrow(resdat))
# coordinates(resdat) <- ~x + y
# rx <- range(resdat$x)
# # rx <- range(aaxis$logarea)
# rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
# # rx[1] <- -6.5  # set temporal interval to -6.5 to have null returns centered
# # ry <- range(resdat$y)
# ry <- range(aaxis$logarea)
# ry <- round(ry / 0.5) * 0.5  # round to nearest 0.5
# 
# # extent data
# # fix area rounded to zero in Jon's plots
# extdat <- data.table("x" = log10(dat$study_duration), 
#                      "y" = log10(dat$sampled_area)) 
# # extdat[is.infinite(x), ]; extdat[is.infinite(y), ]
# 
# extdat$n <- rep(1, nrow(extdat))
# coordinates(extdat) <- ~x + y
# # ex <- range(extdat$x)
# ex <- range(taxis2$logdays)
# ex <- round(ex / 0.5) * 0.5  # round to nearest 0.5
# # ex[1] <- -5
# ey <- range(extdat$y)
# ey <- round(ey / 0.5) * 0.5  # round to nearest 0.5

# density rasters
# vary the kernel size about them
kernsl <- lapply(c(0.4, 0.6, 0.8, 1), function(x) {
  rr <- kdensity(rx[1], rx[2], ry[1], ry[2], 0.1, resdat3, x)
  rr <- (rr / cellStats(rr, sum)) * 100
  er <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat3, x)
  er <- (er / cellStats(er, sum)) * 100
  kerns <- list("res" = rr, "ext" = er)
})

# As separate plots
pdf("paper/figures/res_v_extent_ksize.pdf", width = 7, height = 12, )
par(mfrow = c(4, 2), oma = c(2, 2, 0, 0), mar = c(4, 4, 4, 4))
lapply(kernsl, function(x) {
  # x <- kernsl[[1]]
  image(x$res, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
        ylab = "")#, ylim = range(aaxis1$logarea))
  axis(1, at = taxis1$logdays[-c(5, 9, 10)], labels = tlab1[-c(5, 9, 10)], 
       las = 2, tcl = -0.2, cex.axis = cxa)
  axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
       cex.axis = cxa)  
  aargs <- list(mgp = c(3, 0.25, 0), #at = brklabs, labels = brklabs, 
                cex.axis = cxa, tcl = -0.1)
  plot(x$res, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, #breaks = brks, 
     col = rev(terrain.colors(30)))

  image(x$ext, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
        ylab = "")
  axis(1, at = taxis2$logdays[-5], labels = taxis2$label[-5], las = 2,
       cex.axis = cxa, tcl = -0.2)
  axis(2, at = aaxis2$logarea, labels = alab2[-2], las = 2, cex.axis = cxa,
       tcl = -0.2)
  aargs <- list(mgp = c(3, 0.25, 0), #at = brklabs2, labels = brklabs2, 
                cex.axis = cxa, tcl = -0.1)
  plot(x$ext, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
       legend.shrink = 0.9, #breaks = brks2, 
       col = rev(terrain.colors(30)))
})
par(xpd = NA)
xx <- grconvertX(0.5, from = "ndc", to = "inches")
ii <- c(0.4, 0.6, 0.8, 1)
yy <- c(0.95, 0.75, 0.5, 0.25)
# yy <- c(-1, -5, -10, -20)
for(i in 1:length(ii)) {
  y <- grconvertY(yy[i], from = "ndc", to = "inches")
  # text(x = xx, y = y, labels = paste("kernel =", ii[i]))
  # par(xpd = NA)
  # mtext(text = paste("kernel =", ii[i]), side = 3, line = yy[i])
}
dev.off()
